Hunting experience and prey variability jointly shape individual foraging specialisation in a predator-prey videogame

### 2022-12-22 manuscript/Chapter2-manuscript.Rmd
update ms version (V2) including results

library(data.table)
library(brms)
library(flextable)
library(officer)

@@

*ajouter la description sur les autres variables de performance*

@@

We tested whether foraging behaviour changed with experience, and compared the success of generalists vs specialists, using a multivariate, double-hierarchical generalized linear model (MDHGLM) [@leeDoubleHierarchicalGeneralized2006;@cleasbyQuantifyingPredictabilityBehaviour2015;@westneatBiologyHiddenResidual2014;@Mitchell.etal2016a;@ODea.etal2022]. We categorized the predators by their experience based on the predicted values from the GAMM (see Figure S1 in the supporting information). We assigned the status of "novice" to any predator-player that had a cumulative experience below 100 matches. For example, a predator player that played 500 matches in total would be labelled as a novice in its first 100 matches. Following this logic, we labelled predators as "intermediate" hunters when they had a cumulative experience between 100 and 299 and matches. For the third and last group, we defined predators as "advanced" hunters when they had a cumulative experience equal or above 300, with a maximum of 499 matches.

@@

where $env \sim N(0,\ I\ \sigma^2_{env})$, $av \sim N(0,\ I\ \sigma^2_{av})$, and $id \sim N(0,\ I\ \sigma^2_{id})$. $b$ is the vector of fixed effects with its incidence matrix $X$. $env$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their incidence matrices $M$, $N$, and $Z$. The variance parameters $\sigma^2_{env}$, $\sigma^2_{av}$, and $\sigma^2_{id}$, correspond to the among environment, among avatar, and among individual variances, respectively. $I$ is the identity matrix, with $I,\ \sigma^2$ indicating that the random effects are independently and identically distributed. The $m$ and $d$ subscripts indicate that an estimate is from the mean or dispersion part of the model, respectively. The first two equations were parametrized with a Gaussian distribution family, while the third equation was parametrized with a beta-binomial distribution family. We parametrized the model to run four MCMC chains with 1000 posterior samples for each parameter. To do so, we ran 2500 iterations with a thinning set to 8, with the first 500 iterations used as warmups. We explain how we specified the priors for the MDHGLM in the supporting information.

@@

Contrary to our predictions, individual predators did not differ in the development of their expertise, as the best model was the one with a global smoother with varying intercepts only (Table S1 in the supporting information). The model detected only a weak relationship between hunting success and experience (i.e. the cumulative amount of matches played), such that players did not substantially optimize their success with increasing experience as we expected (Table 1 and Figure S1 in the supporting information). However, predators did differ in their average hunting success across the study period (Figure S1). *peut être coder un autre modèle où je contrôle pour les proies afin de voir si les différences de succès disparaissent*

@@

tab1 <- readRDS(here::here("outputs", "05_outputs_tables", "05_table2.rds"))

set_caption(
  tab1,
  caption = "Posterior mean values, standard errors, and 95% credible intervals for the GAMM with a common global smoother. The values are presented on the logit scale.",
  style = "Table Caption"
)

@@

@@

```{r fig1, out.width = "100%", fig.cap = "**Figure 1.** Coefficient of variation for the predator's speed and for the speed of the prey encountered at each level of experience. (A) Coefficient of variation for the among individual differences in average behaviour. (B) Coefficient fo variation for the among individual differences in intra individual variance (i.e. specialization)."}

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure1.png"))

blablabla figure 2

```{r fig2, out.width = "100%", fig.cap = "**Figure 2.** Distribution of the intra individual behavioural variation (i.e. specialization) of predators when they were (A) novices and (B) advanced. The red band displays the 95% CI of the intercept for the intra individual variance in predator speed."}

@@

```{r fig3, out.width = "100%", fig.cap = "**Figure 3.** Full correlation matrix of the MDHGLM relating the mean and IIV speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Starting from the upper part of the plot are displayed the five parameters for novice predators (gray), the five for intermediate predators (yellow), and the five for advanced predators (blue). Larger dark blue circles indicate stronger positive correlations, while larger dark red circles indicate stronger negative correlations. The black squares indicate correlations between different experience levels."}

@@

*relationship between IIV of prey and success fades with experience*

```{r fig4, out.width = "100%", fig.cap = "**Figure 4.** Title (A) and (B) Novices. (C) and (D) Advanced."}

@@

>>>>

library(data.table)
library(brms)
library(flextable)
library(officer)
library(tidyr)

@@

@@

We tested whether foraging behaviour changed with experience, and compared the success of generalists vs specialists, using a multivariate, double-hierarchical generalized linear model (MDHGLM) [@leeDoubleHierarchicalGeneralized2006;@cleasbyQuantifyingPredictabilityBehaviour2015;@westneatBiologyHiddenResidual2014;@Mitchell.etal2016a;@ODea.etal2022]. We categorized the predators by their experience based on the number of matches they played (i.e. their cumulative experience). We assigned the status of "novice" to any predator-player that had a cumulative experience below 100 matches. For example, a predator player that played 500 matches in total would be labelled as a novice in its first 100 matches. Following this logic, we labelled predators as "intermediate" hunters when they had a cumulative experience between 100 and 299 and matches. For the third and last group, we defined predators as "advanced" hunters when they had a cumulative experience equal or above 300, with a maximum of 499 cumulated matches.

@@

where $env \sim N(0,\ I\ \sigma^2_{env})$, $av \sim N(0,\ I\ \sigma^2_{av})$, and $id \sim N(0,\ I\ \sigma^2_{id})$. $b$ is the vector of fixed effects with its incidence matrix $X$. $env$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their incidence matrices $M$, $N$, and $Z$. The variance parameters $\sigma^2_{env}$, $\sigma^2_{av}$, and $\sigma^2_{id}$, correspond to the among environment, among avatar, and among individual variances, respectively. $I$ is the identity matrix, with $I\sigma^2$ indicating that the random effects are independently and identically distributed. The $m$ and $d$ subscripts indicate that an estimate is from the mean or dispersion part of the model, respectively. Equations (1) and (2) were parametrized with a Gaussian distribution family (predator and prey speed), while equation (3) was parametrized with a beta-binomial distribution family (predator hunting success). We parametrized the model to run four MCMC chains with 1000 posterior samples for each parameter. To do so, we ran 2500 iterations with a thinning set to 8, with the first 500 iterations used as warmups. We explain how we specified the priors of the MDHGLM in the supporting information.

@@

Our results confirm that individual predators displayed important differences in the development of their expertise, as the best model was the one with group-level smoothers exclusively (Table I). For instance, some predators steadily increased their hunting success with experience, while others maintained either a high or low success across their experience (Figure 1A). Moreover, some individuals show a decrease in hunting success with experience. Another pattern that emerged was that some individuals optimized their success at an experience level that was below what we considered as "advanced" (i.e. >300 matches played).

@@

tab1 <- readRDS(here::here("outputs", "05_outputs_tables", "05_table1.rds"))

tab1 <- autofit(tab1)

tab1 <- tab1 %>% 
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      "* 'elpd' refers to the 'expected log pointwise density' and is the value chosen to select the best model"),
    ref_symbols = " "
  ) %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all")

set_caption(
  tab1,
  caption = "Leave-one-out cross-validation table of the three hierarchical GAMMs relating hunting success to player experience.",
  style = "Table Caption"
)

@@

Paragraphe qui parle des différences entre figure 1 A et B, soit, lorsqu'on contrôle pour l'effet des proies.

@@

Contrary to our predictions, we detected only minimal differences in movement speed among individual predators, and these differences did not change with experience (Figure 2A). However, individuals differed in their degree of flexibility, and these differences tended to increase, yet marginally, with experience (Figure 2B). Overall, individual predators appeared to have encountered prey with similar average speeds (Figure 2A), and our results do not support our prediction that predators would encounter prey with varying levels of speed (Figure 2B).

```{r fig2, out.width = "100%", fig.cap = "**Figure 2.** Coefficient of variation for the predator's speed and for the speed of the prey encountered at each level of experience. (A) Coefficient of variation of the among individual differences in average behaviour. (B) Coefficient fo variation of the among individual differences in intra individual variance (i.e. specialization)."}

@@

We predicted that, with experience, predators would become either increasingly flexible, or, more specialized. Here, we found that predators became generally more flexible with experience, but this increase in flexibility was only minimal (Figure 3 A and B). Thus we cannot conclude that experience had an important effect on behavioural specialization. Yet, as we have outlined in Figure 2, it appears that among individual differences in specialization increase when individuals are advanced compared to when they are novice (Figure 3 A and B).

```{r fig3, out.width = "100%", fig.cap = "**Figure 3.** Distribution of the intra individual behavioural variation (i.e. specialization) of predators when they were (A) novices and then (B) advanced hunters. The red band displays the 95% CI of the intercept for the intra individual standard deviation in predator speed."}

@@

At all experience levels, predators that were on average faster were also travelling faster in all their matches (i.e. they were more specialized at travelling fast). Thus, slower predators were also more flexible in their movement through time (Figure 4). This pattern became stronger as predators gained experience (Figure 4). In general, predators tended to be faster and also more specialized in this strategy when they encountered prey that were on average faster (Figure 4). Moreover, predators favored and specialized in a faster foraging mode when encounters with prey were more predictable, and these relationships also became stronger with experience (Figure 4).

The foraging strategies of predators, that is, the correlations between the mean and variance of speed, remained structurally stable among the levels of experience. For instance, novice predators that tended to be faster also kept this strategy as they gained experience (Figure 4 inside the black squares). Overall, we detected the same patterns for within-experience state correlations. The correlations between trait values were also stronger among novice-intermediate and intermediate-advanced states than between novice-advanced states, indicating that behavioural changes in predators occurred as they gained experience.

```{r fig4, out.width = "100%", fig.cap = "**Figure 4.** Full correlation matrix of the MDHGLM relating the mean and IIV speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Starting from the upper part of the plot are displayed the five parameters for novice predators (gray), the five for intermediate predators (yellow), and the five for advanced predators (blue). Larger dark blue circles indicate stronger positive correlations, while larger dark red circles indicate stronger negative correlations. The black squares indicate correlations between different experience levels (i.e. character states)."}

@@

We did not detect significant relationships between the predators' average nor intra individual variation in speed and hunting success (Figure 4), indicating that both slow/flexible and fast/specialist predators were both equally successful. However, when predators were novice, there was a strong positive correlation between the variation in prey that they encountered and their hunting success. Hence, novice individuals that experienced more unpredictable prey encounters at this stage were also more successful. This relationship was decoupled at the later experience levels, such that unpredictable encounters at intermediate and advanced stages were not associated with hunting success anymore.
